---
title: "Conditional Effects by cond_effect()"
author: "Shu Fai Cheung"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Conditional Effects by cond_effect()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width  = 6,
  fig.height = 4,
  fig.align = 'center'
)
```

# What `cond_effect()` Can Do

It can compute the conditional effect of a predictor (the focal variable)
on an outcome variable (dependent variable)
for selected levels of the moderator:

```{r echo = FALSE}
library(stdmod)
data(sleep_emo_con)
lm_out <- lm(sleep_duration ~ age + gender +
                              emotional_stability * conscientiousness,
                              sleep_emo_con)
cond_out <- cond_effect(output = lm_out,
                        x = "emotional_stability",
                        w = "conscientiousness")
print(cond_out, table_only = TRUE)
```

It can also compute standardized conditional moderation effect of this predictor:

```{r echo = FALSE}
data(sleep_emo_con)
lm_out <- lm(sleep_duration ~ age + gender +
                              emotional_stability * conscientiousness,
                              sleep_emo_con)
lm_std <- std_selected(lm_out,
                       to_center = ~ .,
                       to_scale = ~ .)
cond_std <- cond_effect(output = lm_std,
                        x = "emotional_stability",
                        w = "conscientiousness")
print(cond_std, table_only = TRUE)
```

Nonparametric bootstrap percentile confidence interval can also be formed
for standardized conditional effect using `cond_effect_boot()`.

`cond_effect()` is not designed to be a versatile tool. It is designed to
be a function "good-enough" for common scenarios. Nevertheless, it can
report some useful information along with the conditional effects,
as demonstrated below.

# Major Arguments

## Regression Output, Predictor (`x`), and Moderator (`w`)

- `output`: The output of `lm()`, `std_selected()`, or `std_selected_boot()`,
          with at least one interaction term. Bootstrap estimates in
          `std_selected_boot()` will be ignored because bootstrapping will be
          done for each level again.

- `x`: The predictor (focal variable).

- `w`: The moderator.

These are the only required arguments. Just setting them can generate the
graph:


```{r}
library(stdmod)
data(sleep_emo_con)
lm_out <- lm(sleep_duration ~ age + gender +
                              emotional_stability * conscientiousness,
                              sleep_emo_con)
cond_out <- cond_effect(output = lm_out,
                        x = "emotional_stability",
                        w = "conscientiousness")
cond_out
```

By default, the print method of `cond_effect()` output prints the
conditional effects, OLS standard errors, *t* statistics, *p*-values,
and significant test results, along with other information such
as the value of each level of the moderator, its distance from the mean,
the percentage of cases equal to or less than this level. The regression
model is also printed. If only the table of effects is needed, call `print()`
and set `table_only` to `TRUE`:

```{r}
print(cond_out, table_only = TRUE)
```

More options in printing the output can be found in the help page of
`print.cond_effect()`.

# Levels of the Moderator

## Numeric Moderators

If the moderator is a numeric variable, then, by default, the conditional effects
for three levels of the moderators will be used: one standard deviation (SD)
to the mean ("Low"), mean ("Medium"), and one SD above mean ("High").

Users can also use percentiles to define "Low", "Medium", and "High" by setting
`w_method` to `"percentile"`. The default are 16th percentile, 50th percentile,
and 84th percentile, which corresponds approximately to one SD below mean,
mean, and one SD above mean, respectively, for a normal distribution.

```{r}
data(sleep_emo_con)
lm_out <- lm(sleep_duration ~ age + gender +
                              emotional_stability * conscientiousness,
                              sleep_emo_con)
cond_out <- cond_effect(output = lm_out,
                        x = "emotional_stability",
                        w = "conscientiousness",
                        w_method = "percentile")
print(cond_out, title = FALSE, model = FALSE)
```

Note that the empirical percentage of cases equal to or less than a level may
not be exactly equal to that for the requested percentile if the number of
cases is small and/or the number of unique values of the moderator is small.

## Categorical Moderators

If the moderator is a categorical variable (a string variable or
  a factor), then the conditional effect of the moderator for each value of
  this categorical moderator will be printed:

```{r}
set.seed(61452)
sleep_emo_con$city <- sample(c("Alpha", "Beta", "Gamma"),
                               nrow(sleep_emo_con), replace = TRUE)
lm_cat <- lm(sleep_duration ~ age + gender + emotional_stability*city,
                              sleep_emo_con)
cond_out <- cond_effect(lm_cat,
                        x = "emotional_stability",
                        w = "city")
print(cond_out, title = FALSE, model = FALSE)
```

# Nonparametric Bootstrap Confidence Intervals

If one or more variables are standardized, the OLS confidence intervals are not
appropriate (Cheung, Cheung, Lau, Hui, & Vong, 2022; Yuan & Chan, 2011).
Users can
call `cond_effect_boot()` to use nonparametric bootstrapping to form the
percentile confidence interval for each conditional effect.

- `conf`: The level of confidence, expressed as a proportion. Default is
          .95, requesting a 95% confidence interval.
- `nboot` The number of bootstrap samples to drawn. Should be at least 2000
          but 5000 is preferable.

```{r}
lm_out <- lm(sleep_duration ~ age + gender +
                              emotional_stability * conscientiousness,
                              sleep_emo_con)
# Standardize all variables and do the moderated regression again
lm_std <- std_selected(lm_out,
                       to_center = ~ .,
                       to_scale = ~ .)
# `nboot` is the sufficient. Set it to at least 2000 in real analysis
cond_std <- cond_effect_boot(output = lm_std,
                             x = "emotional_stability",
                             w = "conscientiousness",
                             nboot = 500)
print(cond_std, model = FALSE, title = FALSE, level_info = FALSE)
```

# Further Information

Please refer to the help page of `cond_effect()` and `cond_effect_boot()`
for other options available, such as defining the number of SDs from
mean to define "Low" and "High", the percentiles to be used, or
using parallel processing to speed up bootstrapping.

# Reference

Cheung, S. F., Cheung, S.-H., Lau, E. Y. Y., Hui, C. H., & Vong, W. N. (2022)
Improving an old way to measure moderation effect in standardized units.
*Health Psychology*, *41*(7), 502-505. https://doi.org/10.1037/hea0001188.

Yuan, K.-H., & Chan, W. (2011). Biases and standard errors of standardized
regression coefficients. *Psychometrika, 76*(4), 670-690.
https://doi.org/10.1007/s11336-011-9224-6
