---
title: "Mean Center and Standardize Selected Variable by std_selected()"
author: "Shu Fai Cheung and David Weng Ngai Vong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mean Center and Standardize Selected Variable by std_selected()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width  = 6,
  fig.height = 6
)
options(width = 132)
```

# Purpose

This document demonstrates how to use `std_selected()` to do mean centering and
rescaling for selected variables in a regression model. A moderated regression
model is used as an example.

# Setup the Environment

```{r setup}
library(stdmod)
library(lm.beta) # For generating the typical standardized solution
```

# Load the Dataset

```{r load_dataset}
data(sleep_emo_con)
head(sleep_emo_con)
```

This data set has 500 cases of data. The variables are sleep duration,
age, gender,  and the scores from two personality scales, emotional stability
and conscientiousness of the IPIP Big Five markers. Please refer to
(citation to be included) for the detail of the data set.


# Moderated Regression

Suppose we are interested in predicting sleep duration by emotional
stability, after controlling for gender and age. However, we suspect that the
effect of emotional stability, if any, may be moderated by conscientiousness.
Therefore, we conduct a moderated regression as follow:

```{r mod_reg}
lm_raw <- lm(sleep_duration ~ age + gender + emotional_stability*conscientiousness, sleep_emo_con)
summary(lm_raw)
```

The results show that conscientiousness significantly moderates the effect of
emotional stability on sleep duration.

This package has a simple function, `plotmod()`, for generating a typical plot
of the moderation effect:

```{r}
plotmod(lm_raw,
        x = "emotional_stability",
        w = "conscientiousness",
        x_label = "Emotional Stability",
        w_label = "Conscientiousness",
        y_label = "Sleep Duration")
```

The function `plotmod()` also prints the conditional effects of the predictor,
emotional stability in this example.

# Mean Center Conscientiousness

To know the effect of emotional stability when conscientiousness is equal to its
mean, we can center conscientiousness by its mean in the data and redo the
moderated regression. Instead of creating the new variable and rerun the
regression, we can pass the `lm()` output to `std_selected()` and specify the
variables to be mean centered:


```{r}
lm_w_centered <- std_selected(lm_raw, to_center = ~ conscientiousness)
summary(lm_w_centered)
plotmod(lm_w_centered,
        x = "emotional_stability",
        w = "conscientiousness",
        x_label = "Emotional Stability",
        w_label = "Conscientiousness",
        y_label = "Sleep Duration")
```

The argument for meaning centering is `to_center`. The variable is specified
in the formula form.


# Mean Center Conscientiousness and Emotional Stability

This example demonstrates centering more than one variable. In the following
model, both emotional stability and conscientiousness are centered. They are
placed after `~` and joined by `+`.

```{r}
lm_xw_centered <- std_selected(lm_raw, to_center = ~ emotional_stability + conscientiousness)
summary(lm_xw_centered)
plotmod(lm_xw_centered,
        x = "emotional_stability",
        w = "conscientiousness",
        x_label = "Emotional Stability",
        w_label = "Conscientiousness",
        y_label = "Sleep Duration")
```

# Standardize Both Conscientiousness and Emotional Stability

To standardize a variable we first mean center it and then scale it by its
standard deviation. Scaling is done by listing the variable on `to_scale`.
The input format is identical to that of `to_center`.


```{r}
lm_xw_std <- std_selected(lm_raw,
                to_center = ~ emotional_stability + conscientiousness,
                to_scale  = ~ emotional_stability + conscientiousness)
summary(lm_xw_std)
plotmod(lm_xw_std,
        x = "emotional_stability",
        w = "conscientiousness",
        x_label = "Emotional Stability",
        w_label = "Conscientiousness",
        y_label = "Sleep Duration")
```

The function `plotmod()` automatically checks whether a variable is
standardized. If yes, it will report this in the plot.

# Standardize Conscientiousness, Emotional Stability, and Sleep Duration

We can also mean center or standardize the dependent variable. We
just add the variable to the right hand side of `~` in `to_center` and
`to_scale` as appropriate.

```{r}
lm_xwy_std <- std_selected(lm_raw,
                to_center = ~ emotional_stability + conscientiousness + sleep_duration,
                to_scale  = ~ emotional_stability + conscientiousness + sleep_duration)
summary(lm_xwy_std)
plotmod(lm_xwy_std,
        x = "emotional_stability",
        w = "conscientiousness",
        x_label = "Emotional Stability",
        w_label = "Conscientiousness",
        y_label = "Sleep Duration")
```

# Standardize All Variables

If we want to standardize all variables, we can use `~ .` as a shortcut. Note
that `std_selected()` will automatically skip categorical variables (i.e., factors or string
variables in the regression model of `lm()`).

```{r}
lm_all_std <- std_selected(lm_raw, to_center = ~ .,
                                   to_scale  = ~ .)
summary(lm_all_std)
plotmod(lm_all_std,
        x = "emotional_stability",
        w = "conscientiousness",
        x_label = "Emotional Stability",
        w_label = "Conscientiousness",
        y_label = "Sleep Duration")
```

# The Usual Standardized Solution

For comparison, this is the results of standardizing all variables, including
the product term and the categorical variable.

```{r}
lm_usual_std <- lm.beta(lm_raw)
summary(lm_usual_std)
```

# Nonparametric Bootstrap Confidence Interval For Betas

It has been shown (e.g., Yuan & Chan, 2011) that the standard errors of
standardized regression coefficients (betas) computed just by rescaling the variables
are biased, and consequently the confidence intervals are also invalid. The
function `std_selected_boot()` is a version of `std_selected()` that also
report the confidence interval of the regression coefficients when rescaling is conducted,
using nonparametric bootstrapping as suggested by Cheung, Cheung, Lau, Hui, & Vong (2022).

We use the same example above that standardizes emotional stability,
conscientiousness, and sleep duration, to illustrate this function.
The argument `nboot` specifies the number of nonparametric bootstrap samples.
The level of confidence set by `conf`. The default is .95, denoting a 95%
confidence. If this is the desired level, this argument can be omitted.

```{r echo = FALSE}
if (file.exists("eg_lm_xwy_std_ci.rds")) {
    lm_xwy_std_ci <- readRDS("eg_lm_xwy_std_ci.rds")
  } else {
    set.seed(58702)
    lm_xwy_std_ci <- std_selected_boot(lm_raw,
        to_center = ~ emotional_stability + conscientiousness + sleep_duration,
        to_scale  = ~ emotional_stability + conscientiousness + sleep_duration,
        nboot = 2000)
    saveRDS(lm_xwy_std_ci, "eg_lm_xwy_std_ci.rds")
  }
```

```{r eval = FALSE}
lm_xwy_std_ci <- std_selected_boot(lm_raw,
        to_center = ~ emotional_stability + conscientiousness + sleep_duration,
        to_scale  = ~ emotional_stability + conscientiousness + sleep_duration,
        nboot = 2000)
```

```{r}
summary(lm_xwy_std_ci)
```

```{r echo = FALSE}
tmp <- summary(lm_xwy_std_ci)$coefficients
```

The standardized moderation effect is `r formatC(tmp["emotional_stability:conscientiousness", "Estimate"], 4, format = "f")`
, and the 95% nonparametric bootstrap percentile confidence interval is 
`r formatC(tmp["emotional_stability:conscientiousness", "CI Lower"], 4, format = "f")` to
`r formatC(tmp["emotional_stability:conscientiousness", "CI Upper"], 4, format = "f")`.

Note: As a side product, the nonparametric bootstrap confidence of the
other coefficients are also reported. They can be used for other variables that
are standardized in the same model, whether they are involved in the moderation or not.

# Reference

Cheung, S. F., Cheung, S.-H., Lau, E. Y. Y., Hui, C. H., & Vong, W. N. (2022)
Improving an old way to measure moderation effect in standardized units.
Advance online publication. *Health Psychology*. https://doi.org/10.1037/hea0001188.

Yuan, K.-H., & Chan, W. (2011). Biases and standard errors of standardized
regression coefficients. *Psychometrika, 76*(4), 670-690. https://doi.org/10.1007/s11336-011-9224-6
