---
title: "Compute the Standardized Moderation Effect by std_selected"
author: "Shu Fai Cheung and David Weng Ngai Vong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compute the Standardized Moderation Effect by std_selected}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width  = 6, 
  fig.height = 6
)
options(width = 132)
```

# Purpose

This document demonstrates how to use `std_selected()` to compute the correct standardized solution of moderated regression.

# Setup the Environment

```{r setup}
library(stdmod) # For computing the standardized moderation effect conveniently
library(visreg) # For visualizing the moderation effect
library(lm.beta) # For generating the typical standardized solution
```

# Load the Dataset

```{r load_dataset}
data(sleep_emo_con)
head(sleep_emo_con)
```

This data set has 500 cases of data. The variables are sleep duration, age, gender,
and the scores from two personality scales, emotional stability and
conscientiousness of the IPIP Big Five markers. Please refer to
(citation to be added) for the detail of the data set.


# Moderated Regression

Suppose we are interested in predicting sleep duration by emotional stability,
after controlling for gender and age. However, we suspect that the effect of
emotional stability, if any, may be moderated by conscientiousness. Therefore,
we conduct a moderated regression as follow:

```{r mod_reg}
lm_out <- lm(sleep_duration ~ age + gender + emotional_stability*conscientiousness, sleep_emo_con)
summary(lm_out)
visreg(lm_out, "emotional_stability", "conscientiousness",
            breaks = 2, overlay = TRUE)
```

The results show that conscientiousness significantly moderates the effect of
emotional stability on sleep duration.

# The Standardized Solution of the Moderated Regression

To get the correct standardized solution of the moderated regression, with the
product term formed *after* standardization, we can use `std_selected()`.

- The first argument is the regression output from `lm()`. 

- The argument `to_center` tells std_selected which variables need to be mean
  centered. 

- The argument `to_scale` tells std_selected which variables need to be rescaled
  by their standard deviations after centering.

If we want to standardize or mean-center all variables, we can use `~ .` as a
shortcut. Note that `std_selected()` will  automatically skip categorical
variables(i.e., factors or string variables in the regression model of `lm()`).

```{r}
lm_stdall <- std_selected(lm_out, to_center = ~ .,
                                   to_scale  = ~ .)
summary(lm_stdall)
```

The coefficient in this solution,  `r round(coef(lm_stdall)["emotional_stability:conscientiousness"], 5)`, can be interpreted as the increase in the standardized effect of `emotional_stability` for each one standard
deviation increase of `conscientiousness`.

The output from `std_selected()` can be passed to other functions that accept the
output of `lm()`. For example, we can use `visreg` to visualize the moderation
effect using the standardized solution.

```{r}
visreg(lm_stdall, "emotional_stability", "conscientiousness", 
            breaks = 2, overlay = TRUE)
```

# The Usual (Incorrect) Standardized Solution

For comparison, this is the results of standardizing all variables, including
the product term and the categorical variable.

```{r}
lm_beta <- lm.beta(lm_out)
summary(lm_beta)
```

The coefficient of the *standardized* product term is
`r round(coef(lm_beta)["emotional_stability:conscientiousness"], 5)`, which
*cannot* be interpreted as the change in the standardized effect of
`emotional_stability` for each one standard deviation increase of `conscientiousness`.

# The Standardized Solution With Nonparametric Bootstrapping Confidence Intervals

It has been shown (e.g., Yuan & Chan, 2011) that the standard errors of
standardized regression coefficients computed just by rescaling the variables
are biased, and consequently the confidence intervals are also invalid. The
function `std_selected_boot()` is a version of `std_selected()` that also
reports the confidence interval of the regression coefficients when rescaling
is conducted.

We use the same example above that standardizes all variables except for
categorical variables to illustrate this function. The argument `nboot`
specifies the number of nonparametric bootstrap samples. Currently,
only the 95% confidence intervals will be reported.

```{r echo = FALSE}
if (file.exists("eg2_lm_xwy_std_ci.rds")) {
    lm_xwy_std_ci <- readRDS("eg2_lm_xwy_std_ci.rds")
  } else {
    set.seed(649017)
    lm_xwy_std_ci <- std_selected_boot(lm_out, to_center = ~ .,
                                              to_scale  = ~ .,
                                              nboot = 2000)
    saveRDS(lm_xwy_std_ci, "eg2_lm_xwy_std_ci.rds")
  }
```

```{r eval = FALSE}
set.seed(649017)
lm_xwy_std_ci <- std_selected_boot(lm_out,
                                   to_center = ~ .,
                                   to_scale  = ~ .,
                                   nboot = 2000)
```

```{r}
summary(lm_xwy_std_ci)
```

```{r echo = FALSE}
tmp <- summary(lm_xwy_std_ci)$coefficients
```

The standardized moderation effect is 
`r formatC(tmp["emotional_stability:conscientiousness", "Estimate"], 4, format = "f")`,
and the 95% nonparametric bootstrapping confidence interval is
`r formatC(tmp["emotional_stability:conscientiousness", "CI Lower"], 4, format = "f")` to
`r formatC(tmp["emotional_stability:conscientiousness", "CI Upper"], 4, format = "f")`.

Note: As a side product, the nonparametric bootstrapping confidence of the other
coefficients are also reported. They can be used for other variables that are
standardized in the same model, whether they are involved in the moderation or not.

# Reference(s)

Yuan, K.-H., & Chan, W. (2011). Biases and standard errors of standardized
regression coefficients. *Psychometrika, 76*(4), 670–690.
https://doi.org/10.1007/s11336-011-9224-6
