---
title: "Standardized Moderation Effect in SEM by stdmod_lavaan()"
author: "Shu Fai Cheung and David Weng Ngai Vong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Standardized Moderation Effect in SEM by stdmod_lavaan()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width  = 6,
  fig.height = 4,
  fig.align = 'center'
)
options(width = 132)
```

# Purpose

This document demonstrates how to use `stdmod_lavaan()` to compute the
standardized moderation effect in a path analytic model fitted by `lavaan::lavaan()`.


# Setup the Environment

```{r setup}
library(stdmod) # For computing the standardized moderation effect conveniently
library(lavaan) # For doing path analysis in lavaan.
```

# Load the Dataset

```{r load_dataset}
data(sleep_emo_con)
head(sleep_emo_con)
```

This data set has 500 cases of data. The variables are sleep duration, age,
gender,  and the scores from two personality scales, emotional stability and
conscientiousness of the IPIP Big Five markers. Please refer to
(citation to be added) for the detail of the data set.


# Compute the Product Term in the Path Analytic Model

Although `lavaan()` supports the colon operator, `:`, in this example,
we manually compute the product term.

```{r pro_term}
sleep_emo_con$emo_con <- 
  sleep_emo_con$emotional_stability * sleep_emo_con$conscientiousness
```

# Specify the Model in SEM and Fit the Model by `lavaan::lavaan()`

After computing the product term of the model, the remaining process is just
following the normal procedures of constructing SEM in R using `lavaan::lavaan()`
(or related functions such as `lavaan::sem()`).

We specify the model by using age, gender, emotional stability,
conscientiousness, and the product term of emotional stability and
conscientiousness to predict the sleep duration. Therefore, we fit the model
with a moderator as follow: 

```{r mod_sem}
mod <- 
  '
  sleep_duration ~ age + gender + emotional_stability + conscientiousness + emo_con
  '

fit <- sem(mod, sleep_emo_con, likelihood = "wishart")

summary(fit)
```

The results show that conscientiousness significantly moderates the effect of
emotional stability on sleep duration.

*Note*: The argument `Likelihood = wishart` is provided as it ensures that the
SEM results are identical to the linear regression results estimated by OLS,
for illustration only. Users are not required to use this argument when they
are conducting fitting their models in their studies.

# Compute the Standardized Moderation Effect

After fitting the SEM model by `lavaan::lavaan()`, we can use the function
`stdmod_lavaan()` to compute the standardized moderation effect. Noted that
the function can be slow if nonparametric bootstrap confidence interval
is requested by setting `boot_ci` to `TRUE`. Therefore, we highly suggest
using parallel processing if possible when nonparametric bootstrapping
confidence interval is to be computed.

If nonparametric bootstrap confidence interval is requested,
`stdmod_lavaan()` accepts arguments to be passed to `boot::boot()`.
In the following example, the arguments `parallel`, `ncpus`, and `seed`
are arguments for `boot::boot()`. For example, `ncpus` specifies the
number of CPU cores that users wish to use for parallel processing.


```{r echo = FALSE}
if (file.exists("egl_out_boot.rds")) {
    out_boot <- readRDS("egl_out_boot.rds")
  } else {
    set.seed(515464) # Set the random seed
    out_boot <- stdmod_lavaan(fit = fit, x = "emotional_stability",
                              y = "sleep_duration",
                              w = "conscientiousness",
                              x_w = "emo_con",
                              boot_ci = TRUE, R = 2000,
                              parallel = "snow", ncpus = 2)
    saveRDS(out_boot, "egl_out_boot.rds")
  }
```

```{r eval = FALSE}
set.seed(515464) # Set the random seed
out_boot <- stdmod_lavaan(fit = fit, x = "emotional_stability",
                          y = "sleep_duration",
                          w = "conscientiousness",
                          x_w = "emo_con",
                          boot_ci = TRUE, R = 2000,
                          parallel = "snow", ncpus = 2)
```

```{r}
out_boot$stdmod
```

The standardized moderation effect is in the element `stdmod()`. In this
example, the standardized moderation effect is 
`r formatC(out_boot$stdmod, 4, format = "f")`.

Nonparametric bootstrap confidence interval is in the element `ci`.

```{r}
out_boot$ci
```

The 95% confidence interval of the standardized moderation effect is
`r formatC(out_boot$ci["lower_bound"], 4, format = "f")` to
`r formatC(out_boot$ci["upper_bound"], 4, format = "f")`.

By default, the original output from `boot::boot` is stored in the
element `boot_out`. Users can use functions that accept the output
of `boot::boot()`. For example, users can use `boot::boot.ci()`
to get the 90% percentile confidence interval as below:

```{r}
boot::boot.ci(out_boot$boot_out, conf = .90, type = "perc")
```

# Identical to the Results from Moderated Regression

For comparison, these are the standardized solution of the same moderated
regression model using the normal regression function `lm()` and then
processed by `std_selected()`:

```{r echo = FALSE}
if (file.exists("egl_lm_xwy_std_ci.rds")) {
    lm_out <- readRDS("egl_lm_out.rds")
    lm_xwy_std_ci <- readRDS("egl_lm_xwy_std_ci.rds")
  } else {
    lm_out <- lm(sleep_duration ~ age + gender + emotional_stability*conscientiousness,
                sleep_emo_con)
    set.seed(514589)
    lm_xwy_std_ci <- std_selected_boot(lm_out,
                      to_center = ~ emotional_stability + conscientiousness + sleep_duration,
                      to_scale  = ~ emotional_stability + conscientiousness + sleep_duration,
                      nboot = 2000)
    saveRDS(lm_out, "egl_lm_out.rds")
    saveRDS(lm_xwy_std_ci, "egl_lm_xwy_std_ci.rds")
  }
```

```{r eval = FALSE}
lm_out <- lm(sleep_duration ~ age + gender + emotional_stability*conscientiousness,
             sleep_emo_con)
set.seed(514589)
lm_xwy_std_ci <- std_selected_boot(lm_out,
                  to_center = ~ emotional_stability + conscientiousness + sleep_duration,
                  to_scale  = ~ emotional_stability + conscientiousness + sleep_duration,
                  nboot = 2000)
```

```{r}
coef(lm_xwy_std_ci)["emotional_stability:conscientiousness"]
```

We can see that the standardized moderation effect is
`r formatC(coef(lm_xwy_std_ci)["emotional_stability:conscientiousness"], 4, format = "f")`,
equal to that computed by `stdmod_lavaan()`.

# Remarks

The function `stdmod_lavaan()` can be used for more complicated path models.
The computation of the standardized moderation effect in a path model depends
only on the three variables involved (`x`, `w`, and `y`). We used a simple
saturated model in this document only for illustration and comparing the
results obtained form OLS moderated regression. 

# Reference(s)

The computation of the standardized moderation effect is based on the simple
formula presented in the following manuscript, using the standard deviations of
the outcome variable, focal variable, and the moderator:

Cheung, S. F., Cheung, S.-H., Lau, E. Y. Y., Hui, C. H., & Vong, W. N. (2022)
Improving an old way to measure moderation effect in standardized units.
Advance online publication. *Health Psychology*. https://doi.org/10.1037/hea0001188.

