---
title: "Standardized Moderation Effect in a Path Model by stdmod_lavaan()"
author: "Shu Fai Cheung and David Weng Ngai Vong"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Standardized Moderation Effect in a Path Model by stdmod_lavaan()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width  = 6,
  fig.height = 4,
  fig.align = "center"
)
```

# Purpose

This document demonstrates how to use `stdmod_lavaan()` from
the package `stdmod` to compute the
standardized moderation effect in a path model fitted by `lavaan::sem()`.

More about this package can be found
in `vignette("stdmod", package = "stdmod")`
or at [https://sfcheung.github.io/stdmod/](https://sfcheung.github.io/stdmod/).

# Setup the Environment

```{r setup}
library(stdmod) # For computing the standardized moderation effect conveniently
library(lavaan) # For doing path analysis in lavaan.
```

# Load the Dataset

```{r load_dataset}
data(test_mod1)
round(head(test_mod1, 3), 3)
```

This test data set has 300 cases, six variables, all continuous.


# Fit the Model by `lavaan::sem()`

The product term can be formed manually or by the colon operator, `:`.
`stdmod_lavaan()` will work in both cases.

This is the model to be tested:

```{r mod_sem}
mod <-
"
med ~ iv + mod + iv:mod + cov1
dv ~ med + cov2
"
fit <- sem(mod, test_mod1, fixed.x = FALSE)
summary(fit)
```

The results show that `mod` significantly moderates the effect of
`iv` on `med`.

# Compute the Standardized Moderation Effect

As in the case of regression, the coefficient of `iv:mod` in
the standardized solution is not the desired standardized coefficient because
it standardizes the product term.

```{r}
standardizedSolution(fit)[3, ]
```

After fitting the path model by `lavaan::lavaan()`, we can use `stdmod_lavaan()`
to compute the standardized moderation effect using the standard deviations
of the focal variable, the moderator, and the outcome variable
[(Cheung, Cheung, Lau, Hui, & Vong, 2022)](https://doi.org/10.1037/hea0001188).

The minimal arguments are:

- `fit`: The output from `lavaan::lavaan()` and its wrappers, such
          as `lavaan::sem()`.
- `x`: The focal variable, the variable with its effect on the
       outcome variable being moderated.
- `y`: The outcome variable.
- `w`: The moderator.
- `x_w`: The product term.

```{r}
fit_iv_mod_std <- stdmod_lavaan(fit = fit,
                                x = "iv",
                                y = "med",
                                w = "mod",
                                x_w = "iv:mod")
fit_iv_mod_std
```

The standardized moderation effect of `mod` on the `iv`-`med` path is
`r formatC(coef(fit_iv_mod_std), 3, format = "f")`.

# Form Bootstrap Confidence Interval

`stdmod_lavaan()` can also be used to form nonparametric bootstrap
confidence interval for the standardized moderation effect.

There are two approaches to do this. First, if bootstrap
confidence intervals was requested when fitting the model,
the stored bootstrap estimates will be used. This is
efficient because there is no need to do bootstrapping
again.

We fit the model again, with bootstrapping:

```{r echo = FALSE}
if (file.exists("egl_lavaan_boot.rds")) {
    fit <- readRDS("egl_lavaan_boot.rds")
  } else {
    fit <- sem(mod, test_mod1, fixed.x = FALSE,
              se = "boot",
              bootstrap = 2000,
              iseed = 987543)
    saveRDS(fit, "egl_lavaan_boot.rds")
  }
```

```{r eval = FALSE}
fit <- sem(mod, test_mod1, fixed.x = FALSE,
           se = "boot",
           bootstrap = 2000,
           iseed = 987543)
```

If bootstrapping has been done when fitting the model,
just adding `boot_ci = TRUE` is enough to request
nonparametric percentile bootstrap confidence interval:

```{r}
fit_iv_mod_std_ci <- stdmod_lavaan(fit = fit,
                                   x = "iv",
                                   y = "med",
                                   w = "mod",
                                   x_w = "iv:mod",
                                   boot_ci = TRUE)
fit_iv_mod_std_ci
```

The 95% confidence interval of the standardized moderation effect is
`r formatC(confint(fit_iv_mod_std_ci)[1], 3, format = "f")` to
`r formatC(confint(fit_iv_mod_std_ci)[2], 3, format = "f")`.

The second approach, not covered here, uses
[`do_boot()`](https://sfcheung.github.io/manymome/articles/do_boot.html)
from
the [`manymome`](https://sfcheung.github.io/manymome/index.html) package.
to generate bootstrap estimates. To use the stored bootstrap
estimates, set `boot_out` to the output of `do_boot()`.
The stored bootstrap estimates will then be used. This method
can be used when non-bootstrapping confidence intervals are
needed when fitting the model.

# Remarks

The function `stdmod_lavaan()` can be used for more complicated path models.
The computation of the standardized moderation effect in a path model depends
only on the standard deviations of the three variables involved
(`x`, `w`, and `y`).

# Reference(s)

The computation of the standardized moderation effect is based on the simple
formula presented in the following manuscript, using the standard deviations of
the outcome variable, focal variable, and the moderator:

Cheung, S. F., Cheung, S.-H., Lau, E. Y. Y., Hui, C. H., & Vong, W. N. (2022)
Improving an old way to measure moderation effect in standardized units.
*Health Psychology*, *41*(7), 502-505. https://doi.org/10.1037/hea0001188.

